---
title: "02_tidy_data-qpcr"
author: "Bryony Allen"
date: "17/02/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Housekeeping Note #
This document is an RMarkdown document and uses relative paths.

-- RMarkdown -- 
When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
The following arguments allow you to select the code chunks and output that appear in the final document
     {   echo=FALSE} hides the code so the code chunk doesn't appear in the final document 
     {   include=FALSE} hides the code and the output 
      
-- Relative paths -- 
    The working directory will be set automatically to the location of this script file. The following script assumes that in this folder there are subfolders called "code", "data", "doc", "figs", "output", "raw", and "reports" 


# NOTES & SOURCES #
https://www.tidyverse.org/articles/2017/12/workflow-vs-script/     <<<< describes project orientated workflow 

* tidyr guidance
http://garrettgman.github.io/tidying/  << format of dataset 
https://cran.r-project.org/web/packages/tidyr/vignettes/tidy-data.html
https://speakerdeck.com/hadley/welcome-to-the-tidyverse?slide=29    <<<< outlines how tidyr and dplyr can be used to clean data 

* dplyr guidance
https://datacarpentry.org/R-genomics/04-dplyr.html
https://sites.google.com/site/simonthelwall/home/r/data_manipulation_with_dplyr 
https://mgimond.github.io/ES218/Week03a.html    <<< using mutate_at(), list() and ifelse() to apply seperate mutate operations across multiple cols 
https://rstudio-pubs-static.s3.amazonaws.com/116317_e6922e81e72e4e3f83995485ce686c14.html#/   <<< creating new variables in R with mutate() and ifelse() 

* tibble 


** Ranavirus viral load calculation 
Leung, W. T., Thomas-Walters, L., Garner, T. W., Balloux, F., Durrant, C., & Price, S. J. (2017). A quantitative-PCR based method to estimate ranavirus viral load following normalisation by reference to an ultraconserved vertebrate target. Journal of virological methods, 249, 147-155.
https://www.researchgate.net/publication/319881735_1_Ranavirus_detection_and_quantification_by_qPCR_2_Vertebrate_pathogen_load_estimation

# Load required packages 
install.packages("") if necessary  

```{r packages, include=FALSE}
library(here)
library(dplyr)
library(tidyr)
library(tibble)
```

# Part 1: Import data 
>>>> For the time being read in the latest PartialDataset.csv 

Read in the .csv file of merged qPCR outputs, created in "01_import_merge-qpcr" script. 

```{r get data, echo=TRUE}     
raw.data <- read.csv('raw/OLD-data/PartialDataset_Mar2019.csv') 

# when you switch to using RStudio projects and the here() package use the following code 
# raw.data <- read.csv(here("data", raw_PartialDataset_Mar2019.csv') 

```

# Part 2: Tidy data 

Now we start tidying...

A) Sort out the typos 
```{r treatment groups, include=T} 
# the Rv-only group has a space after in $Treatment 
levels(raw.data$Treatment) <- gsub("^Rv $","Rv", levels(raw.data$Treatment))

```

B) Convert the endpoint infection status from "Positive" or "Negative" to a binary form where 0 is negative (*this will change >> in the future you will have to create an endpoint status column for each pathogen and make a loop to assess if the Qmean value lies over the threshold value for that pathogen 

```{r endpoint status, include=T} 
str(raw.data)   # quick check of data type for each column 
  # hmmm at the moment there are more levels to each endpoint.status than if it was just Positive or Negative   

# check and change the "Undetermined" records 
unique(raw.data$Rv.MCPendpoint.status)   #there is an undetermined ....
  raw.data[which(raw.data$Rv.MCPendpoint.status=="Undetermined"),c("ID", "ExperimentNo", "Rv.MCPendpoint.status", "Rv.MCPendpoint.Qmean", "Rv.EBF3Nendpoint.status", "Rv.EBF3Nendpoint.Qmean")]  # it has a Rv QMean score and EBF3N result so change to postive 
    raw.data$Rv.MCPendpoint.status[raw.data$Rv.MCPendpoint.status=="Undetermined"]<-"Positive"
  
unique(raw.data$Rv.EBF3Nendpoint.status)   #there is an undetermined ....
  raw.data[which(raw.data$Rv.EBF3Nendpoint.status=="Undetermined"),c("ID", "ExperimentNo", "Rv.MCPendpoint.status", "Rv.MCPendpoint.Qmean", "Rv.EBF3Nendpoint.status", "Rv.EBF3Nendpoint.Qmean")] 
    raw.data$Rv.EBF3Nendpoint.status[raw.data$Rv.EBF3Nendpoint.status=="Undetermined"]<-"Positive "  # had to put a space here so that it would accept it 

binary.status <-  raw.data %>%    # convert infection status from positive/negaitve to binary 
  mutate(ExperimentNo = as.factor(ExperimentNo)) %>%   # makes experiment number a factor 
  mutate_at(vars(matches("endpoint.status")), list(~ifelse(grepl("Positive", .),1,0)))   # takes seperate ifelse and mutate operations and makes them a single mutate_at() operation 
# selects the columns that have "endpoint.status" in the col name, applies an ifelse operation that finds an value that matches "Positive" (a way round the accidental spaces) and makes then 1's and all other values 0's
# the . acts as a placeholder for each column listed in the list() function and the ~ operatir is used in lieu of .=

glimpse(binary.status)   # another check
```

C) Convert Bd qunatity mean to Genomic Equivalent (GE) by times by 10 which adjusts for the 1:10 dilution ....  note that GE is not calculated for EMA/WS samples because the samples were not diluted (1:10) before qPCR 

```{r Bd clean, include= T}

Bd.clean <- binary.status %>%
   mutate(Bd.endpoint.GE = Bd.endpoint.Qmean*10)  %>%    # creates a new column containing the calculated genomic equivalence score (GE) from QMean (time by 10)
          mutate(viable.GE = EMA.GE.EMA) %>%   # change the name 
           mutate(dead.GE = EMA.GE.WS-EMA.GE.EMA)   # work out the proportion of dead zoospores 

glimpse(Bd.clean)   # check again
```


D) Normalise ranavirus viral load using the calculation as per Leung et al. (2017) 

••• N.B three samples are missing EBF3N qPCR data so for the tie being I have averaged the EBF3N scores and used that for these three samples 

```{r ranavirus normalisation, include=T} 

Bd.clean %>%    # check the Ranavirus positive samples 
  filter(Rv.MCPendpoint.status=='1' & !Treatment=='Bd')  %>%   # only samples that scored positve for MCP (Rv infection) and excluding Bd only treatment group  
    select(ID, Species,ExperimentNo,Treatment,Rv.MCPendpoint.status,Rv.MCPendpoint.Qmean,Rv.EBF3Nendpoint.status,Rv.EBF3Nendpoint.Qmean)

# N.B. three samples are missing EBF3N qPCR outputs  << for now I will manually change them to the mean EBF3N quantity mean score 
    mean(binary.status$Rv.EBF3Nendpoint.Qmean, na.rm=T)  # calculate the mean EBF3N qPCR quantity mean  ==  2592.907

    Bd.clean[which(Bd.clean$Rv.MCPendpoint.status==1 & is.na(Bd.clean$Rv.EBF3Nendpoint.Qmean)),c("ID", "ExperimentNo", "Rv.MCPendpoint.status", "Rv.MCPendpoint.Qmean", "Rv.EBF3Nendpoint.status", "Rv.EBF3Nendpoint.Qmean")]  # finds individuals that have Rv infection (aka MCP score) but haven't got EBF3N data 
    Bd.clean$Rv.EBF3Nendpoint.Qmean[Bd.clean$Rv.MCPendpoint.status==1 & is.na(Bd.clean$Rv.EBF3Nendpoint.Qmean)]<- 2592.907

    
ploidy <- 2   # set the ploidy for the species to 2 

pathogen.clean <- Bd.clean %>%
      mutate(Rv.endpoint.load = Rv.MCPendpoint.Qmean*ploidy / Rv.EBF3Nendpoint.Qmean)   # new column calculating the viral load 

glimpse(pathogen.clean)
```


# Part 2: output cleaned datafile 

```{r get data, include=FALSE}     

write.csv(pathogen.clean,file="data/02_clean-data.csv", row.names=FALSE)

```



################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ###################
