---
title: "03_exploratory-analysis_shedding"
author: "Bryony Allen"
date: "10/05/2019"
output: word_document
---

```{r packages, include=FALSE, warning=FALSE, error=FALSE, message=FALSE}
# Install & load in required packages 
# for tidying data  
library(dplyr)
library(tidyr)
library(tibble)

library(binom) 
library(reshape2)

# for plotting 
library(ggplot2)
library(ggridges)
library(lattice)
library(gcookbook)
```

# Part 1: Import data 

At the moment I only have data on Bd+ reference animals. The Bd+ ref tadpoles received one less dose than the tadpoles in the experiments (Bd x 5, not 6) but otherwise conditions were identical. Before euthanaisa the tadpoles underwent a shedding/filtering trial. See the table below for details


############## update the table below with the correct Bd zoospore counts & the df to have all records ###########

Experiment No. | Species              |  No. of samples  | Total dose (zsp) | Filter membrane 
---------- |------------------------- | -------------| -------------|------------- 
1 | *Bufo bufo* I          | 10      | 3,675,000    |  0.45µm*
2 | *Bufo bufo* II         | 10       | 1,443,750    | 0.45µm
3 | *Rana temporaria*      | 10      | 2,336,250    |
4 | *Alytes muletensis* I  | NA       | 472,500     | (0.2µm)
5 | *Alytes muletensis* II | NA       | 294,759     |  (0.2µm)

* not a filter membrane 

> **N.B.** we didn't have enough animals to have reference individauls for the *Alytes muletensis* experiments 

```{r import data, include=FALSE}

data.shedding <- read.csv('data/01_Bdref-data.csv') 

glimpse(data.shedding)   # check the dataset 
 
data.shedding$ExperimentNo <-  as.factor(data.shedding$ExperimentNo)   # makes experiment number a factor 

data.shedding <- data.shedding %>%
   mutate(Bd.endpoint.GE = Bd.endpoint.Qmean*10) %>%
     mutate(Bd.dpi.1.GE = Bd.dpi.1*10) 
  
```


# Part 2: Summarise the data 

A summary table of the proportion of infected individuals by treatment group and species. (N.B. N shows the number of individuals in that treatment group)

```{r proportion infected, echo=False}
Bd.prop.sum <- data.shedding %>%      # seperate table for Bd
  filter(ExperimentNo=="1") %>%
  group_by(Treatment) %>%    
    summarize(proportion.infected = mean(Bd.endpoint.status, na.rm = TRUE), N=length(Bd.endpoint.status))

Bd.prop.SE <- binom.confint(Bd.prop.sum$proportion.infected * Bd.prop.sum$N, Bd.prop.sum$N, conf.level = 0.95, methods= "logit", Bd.prop.sum)   # calculate SE 
Bd.prop.sum <- bind_cols(Bd.prop.sum,Bd.prop.SE)    # bind to the dataframe

```

Calculating the summary statistics 


# Part 3: Visualise 

```{r plot labels, include=FALSE}
# make label vectors to use in plots
sp.labs <- c(Bb = "Bufo bufo", Rt = "Rana temporaria", Am = "Alytes muletensis")
exp.labs <- c("1" = "Bufo bufo I", "2" = "Bufo bufo II", "3" = "Rana temporaria", "4" = "Alytes muletensis","5" = "Alytes muletensis II")
exp.short.labs <- c("1" = "Bb1", "2" = "Bb2", "3" = "Rt", "4" = "Am")
infect.labs <- c(Bd = "Bd infection", Rv = "Rv infection")
expos.labs <- c("1" = "Bd", "2" = "Rv", "3" = "Bd-Rv", "4"="Rv-Bd")

```

## Part 3a: Bd load distribution 

Look at the distribution of Bd load (GE) for endpoint samples of tadpoles mouthparts.

```{r plot by species, echo=False}

data.shedding %>%     # ridge plot to see distribution of data  
    filter((Bd.endpoint.status=='1' & Bd.endpoint.GE > 0.1))  %>%  ggplot(aes(Bd.endpoint.GE, Treatment)) +
      geom_density_ridges(
                      jittered_points = TRUE, quantile_lines = TRUE, scale = 0.9, alpha = 0.7,
                      vline_size = 1, vline_color = "red",
                      point_size = 0.4, point_alpha = 1,
                      position = position_raincloud(adjust_vlines = TRUE)
          ) + theme_ridges()

```


...and the distribution of Bd load (GE) from the shedding/filtering samples.   

```{r plot by species, echo=False}
data.shedding %>%     # ridge plot to see distribution of data  
    filter((Bd.dpi.1.GE > 0.1))  %>%  ggplot(aes(Bd.dpi.1.GE, Treatment)) +
      geom_density_ridges(
                      jittered_points = TRUE, quantile_lines = TRUE, scale = 0.9, alpha = 0.7,
                      vline_size = 1, vline_color = "red",
                      point_size = 0.4, point_alpha = 1,
                      position = position_raincloud(adjust_vlines = TRUE)
          ) + theme_ridges()

```

The GE scores for the shedding/filtering samples are incredibily low, only just passing the threshold. 


## Part 3b: Bd load comparison 

Plot of individuals endpoint infection load against endpoint shedding load, both reported as Genomic Equivalent (GE) scores above the 0.1 threshold.  

??? **Question** work out whether to log the data or axis ???? 


```{r relationship plot}

data.shedding %>% 
  filter(Bd.endpoint.GE > 0.1 & Bd.dpi.1.GE > 0.1)  %>%
      ggplot(aes(x=Bd.dpi.1.GE, y= Bd.endpoint.GE)) + 
      geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
        theme_bw()


```

Right.... that is not the trend I expected 

...might leave it there for now

################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ###################


```{r full proportion infected, echo=False}
Bd.prop.sum <- data.shedding %>%      # seperate table for Bd
  group_by(ExperimentNo, Treatment) %>%    
    summarize(proportion.infected = mean(Bd.endpoint.status, na.rm = TRUE), N=length(Bd.endpoint.status))

Bd.prop.SE <- binom.confint(Bd.prop.sum$proportion.infected * Bd.prop.sum$N, Bd.prop.sum$N, conf.level = 0.95, methods= "logit", Bd.prop.sum)   # calculate SE 
Bd.prop.sum <- bind_cols(Bd.prop.sum,Bd.prop.SE)    # bind to the dataframe
```