---
title: "03_model-analysis_infection-endpoint_load"
author: "Bryony Allen"
date: "11/04/2019"
output:
  pdf_document:
    toc: false 
    number_sections: true
    fig_caption: yes
    keep_tex: yes
---
# NOTES & SOURCES #
 
Throughout this script I use ExperimentNo" as a proxy for species, where 
"1" = "Bufo bufo I", "2" = "Bufo bufo II", "3" = "Rana temporaria", "4" = "Alytes muletensis" 

Can pathogen load (Bd GE or ranavirus viral load) be explained by pathogen exposure scenario? 
    Do coinfection treatment groups have higher pathogen loads? 
  
Does pathogen load (Bd GE or ranavirus viral load) vary with host species?
Does pathogen load (Bd GE or ranavirus viral load) vary predictably for different levels of pathogen exposure & species?


# Install & load in required packages 

```{r packages, include=FALSE, warning=FALSE, error=FALSE, message=FALSE}
# for tidying data  
library(dplyr)
# library(tidyr)
# library(tibble)

# for plotting 
library(ggplot2)
library(lattice)

# for stats 
# library(binom)
# library(purr)  # purr guidance > for evaluating models http://ijlyttle.github.io/isugg_purrr/presentation.html#(1) 
# library(broom)   # broom summarizes key information about models in tidy tibble()s >> https://github.com/tidymodels/broom
```

# Part 1: Import data 

> At the moment this code chuck reads in a .csv file with endpoint infection data, originally compiled in excel and then cleaned in "02_tidy_data-qpcr"script. In the future you will read in the .csv file of merged qPCR outputs (created, checked and cleaned in "02_tidy_data-qpcr"script) and experiment metadata (checked and cleaned in "02_tidy_data-metadata"script). 

```{r import data, include=FALSE, warning=FALSE, error=FALSE, message=FALSE}

data.endpoint <- read.csv('data/02_clean-data.csv') 

glimpse(data.endpoint)   # check how the dataset has imported (aka whether it has the data type right)  
    data.endpoint$ExperimentNo <-  as.factor(data.endpoint$ExperimentNo)   # convert $ExperimentNo. to factor 
```

# Part 2: Endpoint Infection Load 

GLM's where 
  response variable = Endpoint status [continuous; ] 
  explanatory variable(s) = Treatment [categorical, levels = 4] & Species [categorical, levels = 5]

## Bd

Here I take only individuals that am infected with Bd, at a level over the detection threshold of 0.1GE

```{r Bd load df, echo=FALSE}
Bd.load <- data.endpoint %>%
  filter((Bd.endpoint.status=='1' & Bd.endpoint.GE > 0.1))  %>% 
  select(ID, Species, ExperimentNo, Scenario, Treatment, Bd.endpoint.status, Bd.endpoint.GE) 

droplevels(Bd.load)
```

### Model Selection:  

Choosing a error structure family 

N.B. "the most appropriate link function is the one whcih produces the minimum residual deviance"

... for the log-linear GLMs = arithmetric means whereas the log -transmformed GLM is the geometric mean < interpretation of effects and predictions need to be considered

```{r Bd error structure checks}

# gaussian distribution with data log transformed  
Bd.load2.1 <- glm(log(Bd.endpoint.GE) ~ Treatment + ExperimentNo, data=Bd.load, family= "gaussian")
summary(Bd.load2.1)   #AIC = 680.91  #Residual deviance:  337.15  on 190  degrees of freedom
plot(Bd.load2.1)

# gaussian distribution with log link  <<< TERRIBLE
Bd.load2.2 <- glm(Bd.endpoint.GE ~ Treatment + ExperimentNo, data=Bd.load,family="gaussian"(link='log'))
summary(Bd.load2.2)   #AIC = 680.91  #Residual deviance:  337.15  on 190  degrees of freedom
plot(Bd.load2.2)

# Gamma distribution
Bd.load2.3 <- glm(Bd.endpoint.GE ~ Treatment + ExperimentNo, data=Bd.load, family=Gamma())
summary(Bd.load2.3)   #AIC = 1832.7  #Residual deviance: 239.21  on 190  degrees of freedom
plot(Bd.load2.3)

# Gamma distribution with log link 
Bd.load2.4 <- glm(Bd.endpoint.GE ~ Treatment + ExperimentNo, data=Bd.load, family="Gamma"(link='log'))
summary(Bd.load2.4)   #AIC = 1829.2  #Residual deviance: 235.62  on 190  degrees of freedom
plot(Bd.load2.4)

```

The diagnostic plots are bad 
- residual vs. fitted shows patterning meaning the variance is non-consistent
- QQ plot is S shaped meaning the errors are not normally distributed 
- scale-location the scatter of the residuals increases with the fitted values (bad)
- residual vs. leverage 


... both the linear regression models violate the expectation/assumption of constant normal vairnace 

**** GAMMA error family??? Gamma regression, useful for highly positively skewed data

### Model Comparison: ANOVA 

Analysis of deviance with ```test='Chi'``` selected because of the binomial error family 
... "an ANOVA testing whether the more complex model is significantly better at capturing the data than the simpler model. If the resulting p-value is sufficiently low (usually less than 0.05), we conclude that the more complex model is significantly better than the simpler model, and thus favor the more complex model. If the p-value is not sufficiently low (usually greater than 0.05), we should favor the simpler model."

ANOVA (one-way) 

y= ß1 + ß2xs + ß3xa   
  ...where  
          - baseline value (control) is ß1  (here we use the single pathogen treatment group)
          - levels (treatments) are xs and xa  (here we use coinfection treatment groups)

```{r Bd model comparison}
Bd.load1 <- glm(Bd.endpoint.GE ~ Treatment * ExperimentNo, data=Bd.load, family=Gamma())
Bd.load2 <- glm(Bd.endpoint.GE ~ Treatment + ExperimentNo, data=Bd.load, family=Gamma())
Bd.load3 <- glm(Bd.endpoint.GE ~ Treatment, data=Bd.load, family=Gamma())
Bd.load4 <- glm(Bd.endpoint.GE ~ ExperimentNo, data=Bd.load, family=Gamma())

# compare the simplest models first 
anova(Bd.load2, Bd.load4, test="Chisq") # compares Trt and Sp to just Sp 
anova(Bd.load2, Bd.load3, test="Chisq") # compares Trt and Sp to just Trt 
# compare more complex model to one with interaction 
anova(Bd.load1, Bd.load2, test="Chisq")

```
ANOVA 1: suggests we should reject the more complex model (Treatment + ExperimentNo) in favour for just the model with ExperimentNo only  (pvalue = 0.4725)
ANOVA 2: suggests we should favour the more complex model (Treatment + ExperimentNo) over the model with Treatment only  (pvalue = < .001)  as adding ExperimentNo did lead to significantly improved fit 
ANOVA 3: shows adding a interaction term between Treatment and ExperimentNo did not significantly imporve fit (p-value = 0.122)

### Model Fit: 

To see the fitted values from a regression object (the values of the dependent variable predicted by the model), access the ```fitted.values``` attribute from a regression object with ````$fitted.values```  Look at https://bookdown.org/ndphillips/YaRrr/linear-regression-with-lm.html

```{r Bd fitted values}
names(Bd.load2)   # look at the components of the glm object

Bd.load$glm.fit <- Bd.load2$fitted.values  # add logisitic fitted values back to the dataframe as a new col

head(Bd.load)   # look at the components of the glm object

plot(x = Bd.load$Bd.endpoint.GE, y = Bd.load$glm.fit,  # plot relationship between true values (x-axis) and glm fitted values (y-axis)
     xlab = "True Values",
     ylab = "Model Fitted Values",
     main = "Regression fits of Bd load")

abline(b = 1, a = 0)   # values should fall around this line!

length(unique(Bd.load$ID))

```

Model doesn't fit the data very well. 

Check the model fit for Bd.load4 <- glm(Bd.endpoint.GE ~ ExperimentNo, data=Bd.load, family=Gamma())

```{r Bd fitted values 2}

Bd.load$sp.glm <- Bd.load4$fitted.values  # add logisitic fitted values back to the dataframe as a new col

head(Bd.load)   # look at the components of the glm object

plot(x = Bd.load$Bd.endpoint.GE, y = Bd.load$sp.glm,  # plot relationship between true values (x-axis) and glm fitted values (y-axis)
     xlab = "True Values",
     ylab = "Model Fitted Values",
     main = "Regression fits of Bd load")

abline(b = 1, a = 0)   # values should fall around this line!
```


### Model Plotting:   
To plot the model you need a range of values for which to produce fitted values. Then use the ```predict()``` function to create the model for all the values. ```predict()``` gives you the predicted values based on your (fitted) linear model, the argument type="response" will give you the predicted probabilities 


```{r plot - bi.glm - species and treatment}
Bd.load2 <- glm(Bd.endpoint.GE ~ Treatment + ExperimentNo, data=Bd.load, family=Gamma())

# create a dataframe of "new" data 
newdat <- expand.grid(ExperimentNo=c("1", "2", "3", "4", "5"),Treatment=c("Bd", "Bd-Rv", "Rv-Bd"))

# predict the value/result of the new data using the glm
newdat <-cbind(newdat, predict(object = Bd.load2,   # the model 
                               newdata=newdat, se=TRUE, type="response", print.matrix=T))  # dataframe of new data 
newdat

expl.var <- c(1:3) # chose the range for the x-axis (Experiment No.)
exp.labs <- c("1" = "Bufo bufo I", "2" = "Bufo bufo II", "3" = "Rana temporaria", "4" = "Alytes muletensis I", "5" = "Alytes muletensis II")

newdat1<- subset(newdat, ExperimentNo== "1")    # need to subset the data so you can plot each seperatly 
newdat2<- subset(newdat, ExperimentNo=="2")
newdat3<- subset(newdat, ExperimentNo=="3")
newdat4<- subset(newdat, ExperimentNo=="4")
newdat5<- subset(newdat, ExperimentNo=="5")


Bd.load.predict <- ggplot(newdat, aes(x= expl.var, y= fit, color=ExperimentNo)) +       # plot model estimates, color= the data you subsetted by
  geom_line(data = newdat1, aes(x= expl.var, y= fit), size=1) +                                   # add lines for subset newdat1
  geom_errorbar(data = newdat1, aes(ymin=fit-se.fit, ymax=fit+se.fit), width=.03, size=1)  +      # error bars for subset newdat1
  geom_line(data = newdat2, aes(x= expl.var, y= fit), size=1) +                                   # add lines for subset newdat2
  geom_errorbar(data = newdat2, aes(ymin=fit-se.fit, ymax=fit+se.fit), width=.03, size=1)  +      # error bars for subset newdat2
  geom_line(data = newdat3, aes(x= expl.var, y= fit), size=1) +                                   # add lines for subset newdat1
  geom_errorbar(data = newdat3, aes(ymin=fit-se.fit, ymax=fit+se.fit), width=.03, size=1)  +      # error bars for subset newdat1
  geom_line(data = newdat4, aes(x= expl.var, y= fit), size=1) +                                   # add lines for subset newdat1
  geom_errorbar(data = newdat4, aes(ymin=fit-se.fit, ymax=fit+se.fit), width=.03, size=1) +       # error bars for subset newdat1
  geom_line(data = newdat5, aes(x= expl.var, y= fit), size=1) +                                   # add lines for subset newdat1
  geom_errorbar(data = newdat5, aes(ymin=fit-se.fit, ymax=fit+se.fit), width=.03, size=1) +       # error bars for subset newdat1
    scale_x_continuous(breaks=seq(1:3),labels=c("Bd", "Bd-Rv", "Rv-Bd")) +   # sets the breaks at 1,2 and 3 which correspond to the label names
  ggtitle("Probability of Bd infection load (GE) predicted by\n glm(Bd.endpoint.status ~ Treatment + ExperimentNo, family = Gamma)") 
      
Bd.load.predict.plot <- Bd.load.predict + 
  # ylab("Bd load predictions\n(fit)") +            # TO DO: would be good to have the axis marking 0 & 1 
 # xlab("Treatment Group") +
  guides(fill=guide_legend(title=NULL))  + 
  theme_bw() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

```

```{r export Bd plot}
# export plot as .png 
png("figs/04_model-analysis_infection-end-load.png", type="cairo", units="in", width= 14, height=8, pointsize= 12, res=1000)
print(Bd.load.predict.plot)
dev.off()

```


```{r Bd infection burden: log lm}
mod.burden.Bd.logLM <- lm(log(Bd.endpoint.GE+1) ~ ExperimentNo * Treatment, data=data.endpoint)
summary(mod.burden.Bd.logLM)
```

```{r}
plot(mod.burden.Bd.logLM)

influence.measures(mod.burden.Bd.LM) # table displaying the influence measures of each point  
```

The diagnostic plots are better but not great  
- residual vs. fitted shows patterning meaning the variance is non-consistent
- QQ plot is S shaped meaning the errors are not normally distributed 
- scale-location the scatter of the residuals increases with the fitted values (bad)
- residual vs. leverage 
















################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ###################


--- B) linear regression of Rv viral load by species and treatment interaction--- 

```{r Rv infection burden: LM}
mod.burden.Rv.LM <- lm(Rv.endpoint.load ~ ExperimentNo * Treatment, data=data.endpoint)
summary(mod.burden.Rv.LM)
```

```{r Rv infection burden: LM2}
plot(mod.burden.Rv.LM)
```


--- linear regression of log (Rv viral load) by species and treatment interaction--- 

The distribution of Rv load is very skewed ... could take the log but there are lots of zeros counts (and log(0) is undefined) so could do log(n+1) 
```{r Rv infection burden: log LM}
mod.burden.Rv.logLM <- lm(log(Rv.endpoint.load+1) ~ ExperimentNo * Treatment, data=data.endpoint)
summary(mod.burden.Rv.logLM)
```

```{r Rv infection burden: log LM plot}
plot(mod.burden.Rv.logLM)
```



# Part 3b: better models of infection burden (endpoint infection load) 

A) glm

```{r Bd infection burden: glm}
modGEend <- glm(Bd.endpoint.GE ~ Treatment * ExperimentNo, data=data.endpoint)
summary(modGEend)
```

```{r}
plot(modGEend)
```
Summary: 
- residual deviance, which should be the same as the residual degrees of freedom but is way larger indicates overdispersion (unexplained variation in the response)

Diagnostic plots:
- residual vs. fitted are skewed 
- QQ plot is S shaped with extreme positive residuals 

B) 

Now fit a simpler model has an with only the main effects for Species and Treatment 

```{r binomial glm 2}
modBi2 <- glm(Bd.endpoint.status ~ Species + Treatment, data=subset(data.endpoint, Treatment!="Rv"), family='binomial')
summary(modBi2)

```

Compare the two models using anova: 
```{r binomial glm anova}

anova(modBi, modBi2, test='Chi') 

```

--- D) model selection --- 




: GE :  Bd genomic equivalence 




# ANOVA of linear regression terms 
look at the terms of the model using the anova function
ANOVA tests how much variation in the response variable is explained by each explanatory variable
aka comparing the variation in ge score explained by treamtent (or another variable) to the total variation in ge score (the null model). 
Look at how much smaller the residuals are for the treatment group model to the null model. Graphically, how much shorter are the red residuals than
the blue residuals
```{r linear model ANOVA}

```

# sum of squares of residuals 

We can get the sums of the squares of these residuals from the two models using the function
resid, and then square them and add them up:





# Part 4: Model Simplification 

```{r model simplification }

drop.scope(modQPois)
modQPois2 <- update(modQPois, . ~ . - Strain:Treatment)
anova(modQPois, modQPois2, test='F')

# F) predictions
df <- expand.grid(Strain=levels(coloniesCN$Strain), 
                  Treatment=levels(coloniesCN$Treatment))
predict(modQPois, newdata=df, type='response')

```






################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ###################

