---
title: "03_model-analysis_infection-endpoint_status"
author: "Bryony Allen"
date: "16/02/2018"
output:
  pdf_document:
    toc: false 
    number_sections: true
    fig_caption: yes
    keep_tex: yes
---
# NOTES & SOURCES #
 
I use ExperimentNo as a proxy for species
where "1" = "Bufo bufo I", "2" = "Bufo bufo II", "3" = "Rana temporaria", "4" = "Alytes muletensis"

# Install & load in required packages 

```{r packages, include=FALSE, warning=FALSE, error=FALSE, message=FALSE}
# for tidying data  
library(dplyr)
# library(tidyr)
# library(tibble)

# for plotting 
library(ggplot2)
library(lattice)

# for stats 
library(binom)
# library(purr)  # purr guidance > for evaluating models http://ijlyttle.github.io/isugg_purrr/presentation.html#(1) 
# library(broom)   # broom summarizes key information about models in tidy tibble()s >> https://github.com/tidymodels/broom
```

# Part 1: Import data 

> At the moment this code chuck reads in a .csv file with endpoint infection data, originally compiled in excel and then cleaned in "02_tidy_data-qpcr"script. In the future you will read in the .csv file of merged qPCR outputs (created, checked and cleaned in "02_tidy_data-qpcr"script) and experiment metadata (checked and cleaned in "02_tidy_data-metadata"script). 

```{r import data, include=FALSE, warning=FALSE, error=FALSE, message=FALSE}

data.endpoint <- read.csv('data/02_clean-data.csv') 

glimpse(data.endpoint)   # check how the dataset has imported (aka whether it has the data type right)  
    data.endpoint$ExperimentNo <-  as.factor(data.endpoint$ExperimentNo)   # convert $ExperimentNo. to factor 
```

# Part 2: Endpoint Infection Load 

GLM's where 
  response variable = Endpoint status [continuous; ] 
  explanatory variable(s) = Treatment  [categorical]   &   Species [categorical]

## Bd

Here I have removed the Rv-only treatment group as they have never been exposued to Bd << ? could use them as a control group? I also removed the Am5 (baby Alytes) as they only have one treatment group which I will analyse seperately.   

```{r Bd status df, echo=FALSE}
Bd.status <- data.endpoint %>%
  filter(!Treatment=="Rv") %>%
  filter(!ExperimentNo=='5')  %>% 
  mutate(ExperimentNo = as.factor(ExperimentNo)) %>% 
  select(ID, Species, ExperimentNo, Scenario, Treatment, Bd.endpoint.status, Bd.endpoint.GE) 

droplevels(Bd.status)
```

### Maximal Model: Species*Treamtent interaction 
Fit the most complex model first (maximal model). 

Look at the estimates of the coefficients using ```summary()```
```{r Bd binomial glm}
Bd.status1 <- glm(Bd.endpoint.status ~ Treatment * ExperimentNo, data=Bd.status, family=binomial)
summary(Bd.status1)
```

```{r Bd binomial glm plot, fig.cap= "Diagnostic plots for GLM looking at Bd endpoint infection status by Treamtent group and Species with interactions"}
par(mfrow=c(2,2), mar=c(3,3,3,1), mgp=c(2,0.8,0))
plot(Bd.status1)
```
The diagnostic plots in Figure \@ref(fig:Bd binomial glm plot) aren't ideal 
- residual vs. fitted shows patterning meaning the variance is non-consistent
- residual vs. leverage also shows patterning suggesting certain data points have strong influence  

### Simplification Model 1: Species + Treamtent  
Now fit a simpler model has an with only the main effects for Species and Treatment 
```{r binomial glm 2}
Bd.status2 <- glm(Bd.endpoint.status ~ Treatment + ExperimentNo, data=Bd.status, family=binomial)
summary(Bd.status2)
```

```{r binomial glm plot 2, fig.cap= "Diagnostic plots for GLM looking at Bd endpoint infection status by Treamtent group and Species (no interaction)"}
par(mfrow=c(2,2), mar=c(3,3,3,1), mgp=c(2,0.8,0))
plot(Bd.status2)
```

### Model Comparison: ANOVA 

Analysis of deviance with ```test='Chi'``` selected because of the binomial error family 
... "an ANOVA testing whether the more complex model is significantly better at capturing the data than the simpler model. If the resulting p-value is sufficiently low (usually less than 0.05), we conclude that the more complex model is significantly better than the simpler model, and thus favor the more complex model. If the p-value is not sufficiently low (usually greater than 0.05), we should favor the simpler model."

```{r binomial glm anova}
Bd.status1 <- glm(Bd.endpoint.status ~ Treatment * ExperimentNo, data=Bd.status, family=binomial)
Bd.status2 <- glm(Bd.endpoint.status ~ Treatment + ExperimentNo, data=Bd.status, family=binomial)
Bd.status3 <- glm(Bd.endpoint.status ~ Treatment, data=Bd.status, family=binomial)
Bd.status4 <- glm(Bd.endpoint.status ~ ExperimentNo, data=Bd.status, family=binomial)

# compare the simplest models first 
anova(Bd.status2, Bd.status4, test="Chisq") # compares Trt and Sp to just Sp 
anova(Bd.status2, Bd.status3, test="Chisq") # compares Trt and Sp to just Trt 
# compare more complex model to one with interaction 
anova(Bd.status1, Bd.status2, test="Chisq")

summary(Bd.status2)
```
ANOVA 1: suggests we should reject the more complex model (Treatment + ExperimentNo) in favour for just the model with ExperimentNo only  (pvalue = 0.48)
ANOVA 2: suggests we should favour the more complex model (Treatment + ExperimentNo) over the model with Treatment only  (pvalue = < .001)  as adding ExperimentNo did lead to significantly improved fit 
ANOVA 1: shows adding a interaction term between Treatment and ExperimentNo did not significantly imporve fit (p-value = 0.5752)

### Model Fit:   
To see the fitted values from a regression object (the values of the dependent variable predicted by the model), access the ```fitted.values``` attribute from a regression object with ````$fitted.values```.

```{r fitted values}
names(Bd.status2)   # look at the components of the glm object

Bd.status$bi.glm <- Bd.status2$fitted.values  # add logisitic fitted values back to the dataframe as a new col

head(Bd.status)   # look at the components of the glm object

```


### Model Plotting:   
To plot the model you need a range of values for which to produce fitted values. Then use the ```predict()``` function to create the model for all the values. ```predict()``` gives you the predicted values based on your (fitted) linear model, the argument type="response" will give you the predicted probabilities 

**N.B** the code isn't working here (09.04.19) 

```{r plot - bi.glm - species and treatment}
Bd.status2 <- glm(Bd.endpoint.status ~ Treatment + ExperimentNo, data=Bd.status, family=binomial)

# create a dataframe of "new" data 
newdat <- expand.grid(ExperimentNo=c("1", "2", "3", "4"),Treatment=c("Bd", "Bd-Rv", "Rv-Bd"))

# predict the value/result of the new data using the glm
newdat <-cbind(newdat, predict(object = Bd.status2,   # the model 
                               newdata=newdat, se=TRUE, type="response", print.matrix=T))  # dataframe of new data 
newdat

expl.var <- c(1:3) # chose the range for the x-axis (Experiment No.)
exp.labs <- c("1" = "Bufo bufo I", "2" = "Bufo bufo II", "3" = "Rana temporaria", "4" = "Alytes muletensis")

newdat1<- subset(newdat, ExperimentNo== "1")    # need to subset the data so you can plot each seperatly 
newdat2<- subset(newdat, ExperimentNo=="2")
newdat3<- subset(newdat, ExperimentNo=="3")
newdat4<- subset(newdat, ExperimentNo=="4")

Bd.status.predict <- ggplot(newdat, aes(x= expl.var, y= fit, color=ExperimentNo)) +       # plot model estimates, color= the data you subsetted by
  geom_line(data = newdat1, aes(x= expl.var, y= fit), size=1) +                                   # add lines for subset newdat1
  geom_errorbar(data = newdat1, aes(ymin=fit-se.fit, ymax=fit+se.fit), width=.03, size=1)  +      # error bars for subset newdat1
  geom_line(data = newdat2, aes(x= expl.var, y= fit), size=1) +                                   # add lines for subset newdat2
  geom_errorbar(data = newdat2, aes(ymin=fit-se.fit, ymax=fit+se.fit), width=.03, size=1)  +      # error bars for subset newdat2
  geom_line(data = newdat3, aes(x= expl.var, y= fit), size=1) +                                   # add lines for subset newdat1
  geom_errorbar(data = newdat3, aes(ymin=fit-se.fit, ymax=fit+se.fit), width=.03, size=1)  +      # error bars for subset newdat1
  geom_line(data = newdat4, aes(x= expl.var, y= fit), size=1) +                                   # add lines for subset newdat1
  geom_errorbar(data = newdat4, aes(ymin=fit-se.fit, ymax=fit+se.fit), width=.03, size=1) +       # error bars for subset newdat1
    scale_x_continuous(breaks=seq(1:3),labels=c("Bd", "Bd-Rv", "Rv-Bd"))    # sets the breaks at 1,2 and 3 which correspond to the label names
      
Bd.status.predict.plot <- Bd.status.predict + 
  ylab("Bd status predictions\n(fit)") +            # TO DO: would be good to have the axis marking 0 & 1 
  xlab("Treatment Group") +
  ggtitle("Probability of Bd infection predicted by\n glm(Bd.endpoint.status ~ Treatment + ExperimentNo, 
    family = binomial)") +
  #guides(fill=guide_legend(title="Species")) +      # TO DO: fix the labelling issues here 
  #scale_fill_discrete(labels=exp.labs) +
  theme_bw() +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

# export plot as .png 
png("figs/04_model-analysis_infection-end-status.png", type="cairo", units="in", width= 14, height=8, pointsize= 12, res=1000)
print(Bd.status.predict.plot)
dev.off()

```

# Part 3a: models of infection burden (endpoint infection load) 

Can pathogen load (Bd GE or ranavirus viral load) be explained by pathogen exposure scenario? 
    Do coinfection treatment groups have higher pathogen loads? 
  
Does pathogen load (Bd GE or ranavirus viral load) vary with host species?
Does pathogen load (Bd GE or ranavirus viral load) vary predictably for different levels of pathogen exposure & species?

--- A) linear regression of Bd genomic equivalence by species and treatment interaction --- 

```{r Bd infection burden: lm}
mod.burden.Bd.LM <- lm(Bd.endpoint.GE ~ ExperimentNo * Treatment, data=data.endpoint)
summary(mod.burden.Bd.LM)
```
- residuals skewed (not symmetrical around 0 and should be similar magnitude)
- 

```{r}
plot(mod.burden.Bd.LM)
```
The diagnostic plots are bad 
- residual vs. fitted shows patterning meaning the variance is non-consistent
- QQ plot is S shaped meaning the errors are not normally distributed 
- scale-location the scatter of the residuals increases with the fitted values (bad)
- residual vs. leverage 


```{r Bd infection burden: log lm}
mod.burden.Bd.logLM <- lm(log(Bd.endpoint.GE+1) ~ ExperimentNo * Treatment, data=data.endpoint)
summary(mod.burden.Bd.logLM)
```

```{r}
plot(mod.burden.Bd.logLM)

influence.measures(mod.burden.Bd.LM) # table displaying the influence measures of each point  
```

The diagnostic plots are better but not great  
- residual vs. fitted shows patterning meaning the variance is non-consistent
- QQ plot is S shaped meaning the errors are not normally distributed 
- scale-location the scatter of the residuals increases with the fitted values (bad)
- residual vs. leverage 


--- B) linear regression of Rv viral load by species and treatment interaction--- 

```{r Rv infection burden: LM}
mod.burden.Rv.LM <- lm(Rv.endpoint.load ~ ExperimentNo * Treatment, data=data.endpoint)
summary(mod.burden.Rv.LM)
```

```{r Rv infection burden: LM2}
plot(mod.burden.Rv.LM)
```


--- linear regression of log (Rv viral load) by species and treatment interaction--- 

The distribution of Rv load is very skewed ... could take the log but there are lots of zeros counts (and log(0) is undefined) so could do log(n+1) 
```{r Rv infection burden: log LM}
mod.burden.Rv.logLM <- lm(log(Rv.endpoint.load+1) ~ ExperimentNo * Treatment, data=data.endpoint)
summary(mod.burden.Rv.logLM)
```

```{r Rv infection burden: log LM plot}
plot(mod.burden.Rv.logLM)
```


... both the linear regression models violate the expectation/assumption of constant normal vairnace 

**** GAMMA error family??? Gamma regression, useful for highly positively skewed data

# Part 3b: better models of infection burden (endpoint infection load) 

A) glm

```{r Bd infection burden: glm}
modGEend <- glm(Bd.endpoint.GE ~ Treatment * ExperimentNo, data=data.endpoint)
summary(modGEend)
```

```{r}
plot(modGEend)
```
Summary: 
- residual deviance, which should be the same as the residual degrees of freedom but is way larger indicates overdispersion (unexplained variation in the response)

Diagnostic plots:
- residual vs. fitted are skewed 
- QQ plot is S shaped with extreme positive residuals 

B) 

Now fit a simpler model has an with only the main effects for Species and Treatment 

```{r binomial glm 2}
modBi2 <- glm(Bd.endpoint.status ~ Species + Treatment, data=subset(data.endpoint, Treatment!="Rv"), family='binomial')
summary(modBi2)

```

Compare the two models using anova: 
```{r binomial glm anova}

anova(modBi, modBi2, test='Chi') 

```




```{r fitted values} 
# https://bookdown.org/ndphillips/YaRrr/linear-regression-with-lm.html
names(Bd.status2)   # look at the components of the glm object

Bd.status$bi.glm <- Bd.status2$fitted.values

# plot the relationship between true values and glm fitted values
plot(x = Bd.status$Bd.endpoint.status,                          # True values on x-axis
     y = Bd.status$bi.glm,               # fitted values on y-axis
     xlab = "True Values",
     ylab = "Model Fitted Values",
     main = "Regression fits of Bd status")

abline(b = 1, a = 0)                             # Values should fall around this line!
```









--- D) model selection --- 

ANOVA (one-way) 

y= ß1 + ß2xs + ß3xa   
  ...where  
          - baseline value (control) is ß1  (here we use the single pathogen treatment group)
          - levels (treatments) are xs and xa  (here we use coinfection treatment groups)


: GE :  Bd genomic equivalence 




# ANOVA of linear regression terms 
look at the terms of the model using the anova function
ANOVA tests how much variation in the response variable is explained by each explanatory variable
aka comparing the variation in ge score explained by treamtent (or another variable) to the total variation in ge score (the null model). 
Look at how much smaller the residuals are for the treatment group model to the null model. Graphically, how much shorter are the red residuals than
the blue residuals
```{r linear model ANOVA}

```

# sum of squares of residuals 

We can get the sums of the squares of these residuals from the two models using the function
resid, and then square them and add them up:





# Part 4: Model Simplification 

```{r model simplification }

drop.scope(modQPois)
modQPois2 <- update(modQPois, . ~ . - Strain:Treatment)
anova(modQPois, modQPois2, test='F')

# F) predictions
df <- expand.grid(Strain=levels(coloniesCN$Strain), 
                  Treatment=levels(coloniesCN$Treatment))
predict(modQPois, newdata=df, type='response')

```


# Part 5: 








################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ################### SCRAPS ###################


data.endpoint %>% 
   filter((Bd.endpoint.status=='1' & Bd.endpoint.GE > 0.1))  %>% 
     do(model = lm(Bd.endpoint.GE ~ Species + Treatment, data = .))

mod.Bd.status.bi <- data.endpoint %>%
  filter(!Treatment== "Rv") %>% do(model = glm(Bd.endpoint.status ~ Treatment * ExperimentNo, data=., family=binomial)) 
  
```{r do models worked example}
by_cyl <- group_by(mtcars, cyl)
do(by_cyl, head(., 2))
models <- by_cyl %>% do(mod = lm(mpg ~ disp, data = .))
models

summarise(models, rsq = summary(mod)$r.squared)
models %>% do(data.frame(coef = coef(.$mod)))
models %>% do(data.frame(
  var = names(coef(.$mod)),
  coef(summary(.$mod)))
)

```


#### 

anova(
)

newdat <- expand.grid(symbiont=c("S-", "S+"),
                      wasp_strain=c("Madeira", "Lh14", "France"),
                      ethanol=c("EtOH-", "EtOH+"))  
newdat <-cbind(newdat, predict(wm5, newdata=newdat, se=TRUE, type="response", print.matrix=T))
newdat #the complete dummy dataset with predicted effect sizes

plot(fit~expl.var, data = data)
